New insights into the Weddell Sea ecosystem applying a quantitative network approach

### 2022-12-29 Manuscript/Manuscript.Rmd
Minor changes & caveat

---
title: New insights into the Weddell Sea ecosystem applying a quantitative network approach
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Tomás I.
    surname: Marina
    affiliation: "1, *"
    email: tomasimarina@gmail.com
  - given_name: Leonardo A.
    surname: Saravia
    affiliation: "1, *"
    email: arysar@gmail.com
  - given_name: Susanne
    surname: Kortsch
    affiliation: 2
    # authors can have multiple affiliations, which can also be used to mark deceased coauthors
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: Spatial Foodweb Ecology Group, Department of Agricultural Sciences, University of Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  Network approaches can shed light on the structure and stability of complex marine communities. In recent years, such approaches have been successfully applied to study polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea is one of the most studied marine ecosystems outside the Antarctic Peninsula in the Southern Ocean. Yet, few studies consider the known complexity of the Weddell Sea food web which in its current form comprises 490 species and 16041 predator-prey interactions. Here we analysed the Weddell Sea food web, focusing on trophic interactions that underpin ecosystem structure and stability. We estimated the strength for each interaction in the food web, characterised species position in the food web using unweighted and weighted properties, and analysed species’ roles with respect to the stability of the food web. On one hand, we found that the distribution of the interaction strength at the food web level is asymmetric, where weak interactions are prevalent. Including such information as a (weighted) property for species we detected a positive relationship between species mean interaction strength and two unweighted properties, trophic level and the total number of interactions. We also found that only a few species are key in terms of food web stability, presenting high mean interaction strength, mid to high trophic level, relatively high number of interactions, and mid to low trophic similarity. In the same analysis we have integrated food web and species information, enabling a more complete assessment of the ecosystem structure and function, likely highlighting the ecological processes at play in the Weddell Sea. We consider that our results provide new insights important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

@@

<!-- LAS I think that the methodology of estimating the IS is in the heart of the paper so we need to say something about that, this is why I added the following paragraph-->
Several methodologies have been applied to estimate the interaction strength in food webs, where the quantity and quality of data mostly determine which approach is more convenient [@Berlow2004]. Experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances trough time [@Fahimipour2014; @Chang2021], and/or estimation of metabolic rates and biomass of all species in the community [@Neutel2014]; all these methods require a large experimental and/or sampling effort that restrict them to simple networks (circa 10 species). Other methods based on predator-prey body mass allometric relationships [@Berlow2009; @Pawar2012] could be applied to larger networks due to less data requirements, although they might have less power with increasingly complex networks unless more traits are included in the estimations [@Jonsson2018]. For food webs composed by nearly 1000 species and about 20000 interactions only those methods with less data requirements are feasible. On of these is the one proposed by @Pawar2012, which allows to obtain interaction strength estimations by combining data on body mass (of consumer -predator- and resource -prey-) and interaction dimensionality (classified as 2 or 3 dimensions).

<!-- Paragraph on unweighted and weighted networks here? -->
We should maybe write a paragraph about weighted and unweighted approaches here, where we aslo introduce the approach we used... @Susanne could you write it?

<!-- Insights of network science to polar ecosystems -->
> [susanne] I moved the paragraph below. it makes more sense to first talk about polar regions in genaral and then introduce the Weddell Sea.
>
> [Tomás] I agree.

Using a network approach, different types (e.g., terrestial, lake, marine) of food webs from various geographic locations are studied worldwide, including marine polar food webs [@Carscallen2012; @Santana2013; @Kortsch2019; @Pecuchet2022]. Some of these studies from the Arctic show how food web properties (e.g., connectance) are constrained by environmental factors such as sea ice cover and seawater temperature [@Kortsch2019; @Pecuchet2022]. In the Southern Ocean, important insights have been gained into mechanisms of energy flow, the relative importance and traits of individual species and the influence of environmental variables (e.g., sea-ice) on the structure of local food webs [@Cordone2020; @Rossi2019]. For instance, in Potter Cove (West Antarctic Peninsula) the substratum type (i.e. hard/soft or rocks/sediments) plays a significant role in the structure and stability of the food web. In Terra Nova Bay (Ross Sea), the architecture of biodiversity was reshaped by the pulsed input of sympagic food sources following sea-ice break up, with food web simplification, decreased intraguild predation, potential disturbance propagation and increased vulnerability to biodiversity loss [@Rossi2019].
> [susanne] what do you mean here: the relative importance and traits of individual species? What important mechanisms? Can we be more specefic? Can we also be more specific about how environmental variables (e.g., sea-ice) influence the structure of local food webs in the SO?
> 
>  [Tomás] I added 2 phrases in this regard.

<!-- Climate change in Weddell Sea -->
The Weddell Sea is expected to be one of the last regions of the Southern Ocean to experience the consequences of climate change due to its extensive ice cover and ocean currents [@Teschke2021] resulting in less sea surface warming compared to other areas of the Southern Ocean. This Sea plays an important role in driving global thermohaline circulation and ventilating the global abyssal ocean because it generates a considerable part of the Antarctic Bottom Water [@Fahrbach2009]. Because of these environmental characteristics, the Weddell Sea may serve as a refuge for Antarctic species which depend on sea ice (e.g. krill, emperor penguin, Weddell seal) or have low heat tolerance (e.g., most notothenioid fishes) due to their adaptations to freezing tempeartures [@Griffiths2017]. While essential large-scale hydrodynamic relationships are relatively well-known for this region [@deSteur2019], information on the current distribution, abundance and sensitivity to climate change is only partially known for a few species (e.g., emperor penguin) [@Houstin2022].
> [susanne]it is not clear from the text why Weddell Sea is one of the last regions of the SO to face the consequences of warming, be explicit here, how are the currents affecting the Weddell Sea?
>
>[Tomás] I added 2 phrases in this regard.
>
> @learsar could you add this reference to Zotero? Fahrbach, E., Beszczynska-Möller, A., & Rohardt, G. (2009). Polar oceans-an oceanographic overview. In Biological Studies in Polar Oceans-Exploration of Life in Icy Waters, G. and H. Hempel (Eds.), Wirtschaftsverlag NW, Bremerhaven (pp. 17-36). https://epic.awi.de/id/eprint/23118/

<!-- Food web studies about Weddell Sea / Key species or species role -->
Overall, the complexity the Weddell Sea food web is high and consists of 488 species and 16200 predator-prey interactions[@Jacob2011]. In an attempt to better understand species role related to food web stability, Jacob et al. (2011) performed secondary extinction experiments and found that the removal of small to medium-sized, and not large, organisms caused a cascade of secondary extinctions. This findings highlighted the relative importance of predators, rather than prey, in the architecture, functioning and stability of the Weddell Sea food web, which coincides with findings from recent meta-analyses in natural complex food webs [@Brose2019; @Perkins2022]. Other investigations have considered this food web though in a meta-analysis context, showing that high predator-prey body-mass ratios, which create weak interactions that stabilize communities against perturbations and maintain ecosystem functioning, are found for predator groups with specific trait combinations including small vertebrates and large swimming or flying predators [@Brose2019].

> [susanne] more text here, WHAT DO THESE META_ANALSYES STUDIES SHOW [@Brose2019; @Perkins2022; @Saravia2022].
>
> [Tomás] I added specific results from Brose2019.

<!-- Objectives -->
In this study, we aim to go beyond a purely topological (presence/absence) assessment of who eats whom in the Weddell Sea ecosystem by providing a quantitative analysis of the trophic interaction network. We aim to analyse the species' role for the structure and stability of the food web. To achieve this, we: 1) estimated the strength for each interaction in the Weddell Sea food web, 2) characterised species role considering both weighted and unweighted properties, and 3) analysed the species’ role realted to the stability of the food web. This is the first time that the interaction strength was estimated for all the interactions mostly at the species level for the Weddell sea food web

@@

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS, stretching approximately 450 km in length from East to West (Figure 1). Water depth varies between 200 and 500 meters, and shallower areas are covered by continental ice, which forms the coastline along the eastern and southern part of the Weddell Sea. The shelf area contains a complex three-dimensional benthic habitat with large benthic biomasses, intermediate to high diversity in comparison to benthic boreal communities and a spatially patchy distribution of organisms [@Dayton1990; @Teixido2002].

@@

To analyse the impact of species on food web stability, we performed extinction simulations deleting one species at a time, that is for every extinction network size was reduced by one species only. After each extinction, we calculated the stability of the network minus the targeted species (489 nodes) and compared it with that of the whole network (490 nodes in total). To calculate stability, we used the mean of the real part of the maximum eigenvalue of the Jacobian matrix using randomized Jacobians and keeping the predator-prey sign structure fixed [@Allesina2008; @Grilli2016]. This stability index indicates a more stable food web when it is negative. We performed 1000 simulations for each species extinction and obtained a mean maximum eigenvalue for each case. At last we statistically analysed such a difference with an Anderson-Darling test considering a p-value < 0.01 [@Scholz1987]. If this difference is positive, then the stability of the food web is higher if a species was removed, and vice versa. A detailed description on the stability calculations can be found in Supplementary Material.

Once we had the results for the impact on stability for each species extinction, we plotted them considering weighted (interaction strength) and unweighted properties, and species habitat. With this we aim to characterise those species with a relatively high effect on the stability of the food web.

@@

We employed a range of descriptors using both unweighted and weighted food web properties to characterise the dynamic and multifaceted nature of the Weddell Sea food web. Our results show a positive relationship between interaction strength and trophic level, and between interaction strength and species degree. Mean interaction strength increases with trophic level and species degree. The former relationship contradicts studies that suggest that mid-trophic level species (which, krill??? examples are always good) are involved in the major pathways of energy flow in high-latitude marine ecosystems [@Pinkerton2014; @Murphy2016; @McCormack2020; @Riccialdelli2020]. However, this finding could also be explained by the lack of species biomass information in the calculation of interaction strength using the methodology we applied [@Parwar2012]. The methodlogy by Parwar et al.(2012) allows information on species density/biomass to be included, but this type of data was not available for the majority of  species in our food web. In the Weddell Sea, species with high degree also tend to have high mean interaction strengthsThe positive relationship between interaction strength and species degree reinforces the central role of species with many interactions: species with a high degree (hubs) have a large impact on overall food web structure and functioning [@Dunne2002a; @Kortsch2015]. This information on the quantity and quality of interactions and its relationship enables a robust assessment of the species’ role in the stability of the food web [@Cirtwill2018a].

@@

Only a few species play a key role with respect to the Weddell Sea food web stability, according to the mean maximum eigenvalue stability index employed in this study. This is in concordance with other studies on complex empirical food webs in marine ecosystems in the Arctic and other locations in Antarctica [@Kortsch2015; @Marina2018; @Rodriguez2022]. These key species are characterised by a particular set of food web properties: high to mean IS; mid to high trophic level; a relatively high number of interactions; and mid to low trophic similarity. In a previous study on sequential extinction simulations for the Weddell Sea food web [@CITA!!!], it was found  that larger bodied-sized species could be lost without causing a collapse of the network. A major caveat of this finding, also recognised by the authors, was that population dynamics were ignored and hence no top-down extinctions, or other indirect effects, could occur. In our study we considered such top-down effects by including information on the species IS, which is of paramount importance when analysing the response of perturbations in ecological communities [@McCann1998; @Montoya2009; @Novak2011]. Thus, our study suggests that species with high mean IS and high trophic levels need to be considered with particular attention when trying to predict the effects of perturbations on the Weddell Sea ecosystem. This conclusion is further reinforced by the finding that these species have mid to low trophic similarity, which means that few other species of the food web can occupy the same trophic role. In a review, it was emphasised that polar pelagic communities are particularly sensitive to changes due to a low functional redundancy at key trophic levels [@Murphy2016]. Here we provide broader analyses of species impact on food web robustness by including species from all habitats (benthic, pelagic and land-based). This suggests that the sensitivity of marine polar ecosystems to environmental perturbations is a concern also beyond the pelagic realm.

@@

We consider that the information provided in this study is important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea [@Teschke2021].

>>>>

---
title: New insights into the Weddell Sea ecosystem applying a quantitative network approach
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Tomás I.
    surname: Marina
    affiliation: "1, *"
    email: tomasimarina@gmail.com
  - given_name: Leonardo A.
    surname: Saravia
    affiliation: "1, *"
    email: arysar@gmail.com
  - given_name: Susanne
    surname: Kortsch
    affiliation: 2
    # authors can have multiple affiliations, which can also be used to mark deceased coauthors
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: Spatial Foodweb Ecology Group, Department of Agricultural Sciences, University of Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  Network approaches can shed light on the structure and stability of complex marine communities. In recent years, such approaches have been successfully applied to study polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea is one of the most studied marine ecosystems outside the Antarctic Peninsula in the Southern Ocean. Yet, few studies consider the known complexity of the Weddell Sea food web, which in its current form comprises 490 species and 16041 predator-prey interactions. Here we analysed the Weddell Sea food web, focusing on the species and trophic interactions that underpin ecosystem structure and stability. We estimated the strength for each interaction in the food web, characterised species position in the food web using unweighted and weighted food web properties, and analysed species’ roles with respect to the stability of the food web. We found that the distribution of the interaction strength at the food web level is asymmetric, with many weak interactions and few strong ones. We detected a positive relationship between species mean interaction strength and two unweighted properties (i.e., trophic level and the total number of interactions). We also found that only a few species possess key positions in terms of food web stability. These species are charcterised by high mean interaction strength, mid to high trophic level, relatively high number of interactions, and mid to low trophic similarity. In this study, we integrated unweighted and weighted food web information, enabling a more complete assessment of the ecosystem structure and function of the Weddell Sea food web. Our results provide new insights, which are important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

@@

<!-- Paragraph on unweighted and weighted networks  -->
Estimating interaction strength in food webs allows differentiating the importance of species interactions. On the contrary, unweighetd food web representations give equal importantance to all interactions well-knowing that some species interactions are stronger than others and hence play a diffrent role for ecosystem functioning and stability. Both empirical and theoretical studies show that interactions strength distributions in food webs are assymetric [@Paine1992; @McCann1998; @Emmerson2004; @Wootton2005; @Kortsch2021], containing a few strong and many weak interactions. This asymmetric patterning of weak and strong links is crucial to food web stability [@Paine1992; @McCann1998; @Neutel2022]. In a recent paper on an aquatic food web it was further highlighted that temporal changes in ecosytem functioning could only be predicted using weighted food web structure [@Pecuchet2022 or @Kortsch2021?]. Hence, in order to assess the stability and functioning of a food web, it is important to first determine the quantitative structure of the trophic network.

<!-- LAS I think that the methodology of estimating the IS is in the heart of the paper so we need to say something about that, this is why I added the following paragraph-->
Several methodologies have been applied to estimate the interaction strength in food webs, where the quantity and quality of the data mostly determines which approach is the most convenient [@Berlow2004]. Approaches include experimental methods combined with dynamic modelling [@Emmerson2004; @Carrara2015], measurements of species abundances through time [@Fahimipour2014; @Chang2021], and estimation of metabolic rates and biomass of all species in the community [@Neutel2014]. However, these types of methods require large experimental set-ups and parameterisations restricting the analyses to smaller networks (e.g., approximatley 10 species or less). Other methods based on allometric scaling relationships and biomass information [@Kortsch 2021; @Gauzens2018] can be applied to larger networks with less data requirements, but this comes at the expense of precision in the predictions. For even larger food webs composed of nearly 1000 species and more than 10000 interactions, only methods with even less data requirements are feasible. One of these methods, proposed by @Pawar2012, combines data on consumer and resource body masses and consumer search space (interaction dimensionality) to obtain interaction strength estimates for each pairwise predator-prey interaction. An advantage of this method is that it can be applied without information on species biomass.[WHAT ARE THE IMPLICATION OF THIS; NOT INCLUSDING BIOMASS; WHAT IS THE ASSUMPTION WE ARE MAKING; MAYBE WE NEED TO ADD A FEW WORDS ABOUT THIS IN THE DISCUSSION]

[Leo; what do you mean with less powerful? in terms of what be explicit, do you mean precise?]

<!-- Insights of network science to polar ecosystems -->
Using a a network approach, different types (e.g., terrestial, lake, marine) of food webs from various geographic locations are studied worldwide, including marine polar food webs [@Carscallen2012; @Santana2013; @Kortsch2019; @Pecuchet2022]. Some of the studies from the Arctic show how food web properties (e.g., connectance) are constrained by environmental factors such as sea ice cover and seawater temperature [@Kortsch2019; @Pecuchet2022]. In the Southern Ocean, important insights have been gained into mechanisms of energy flow, the relative importance of individual species and their traits, and the influence of environmental variables (e.g., sea-ice) on the structure of local food webs [@Cordone2020; @Rossi2019]. For instance, in Potter Cove (West Antarctic Peninsula) the substratum type (i.e. hard/soft or rocks/sediments) plays a significant role in the structure and stability of the food web. In Terra Nova Bay (Ross Sea), the architecture of biodiversity was reshaped by the pulsed input of sympagic food sources following sea-ice break up, with food web simplification, decreased intraguild predation, potential disturbance propagation and increased vulnerability to biodiversity loss [@Rossi2019].
> [susanne] what do you mean here: the relative importance and traits of individual species? What important mechanisms? Can we be more specefic? Can we also be more specific about how environmental variables (e.g., sea-ice) influence the structure of local food webs in the SO?
> 
>  [Tomás] I added 2 phrases in this regard.

<!-- Climate change in Weddell Sea -->
The Weddell Sea is expected to be one of the last regions of the Southern Ocean to experience the consequences of climate change due to its extensive ice cover and ocean currents [@Teschke2021] resulting in less sea surface warming compared to other areas of the Southern Ocean. This Sea plays an important role in driving global thermohaline circulation and ventilating the global abyssal ocean because it generates a considerable part of the Antarctic Bottom Water [@Fahrbach2009]. Because of these environmental characteristics, the Weddell Sea may serve as a refuge for Antarctic species which depend on sea ice (e.g. krill, emperor penguin, Weddell seal) or have low heat tolerance (e.g., most notothenioid fishes) due to their adaptations to freezing tempeartures [@Griffiths2017]. While essential large-scale hydrodynamic relationships are relatively well-known for this region [@deSteur2019], information on the current distribution, abundance and sensitivity to climate change is only partially known for a few species (e.g., emperor penguin) [@Houstin2022].

> @learsar could you add this reference to Zotero? Fahrbach, E., Beszczynska-Möller, A., & Rohardt, G. (2009). Polar oceans-an oceanographic overview. In Biological Studies in Polar Oceans-Exploration of Life in Icy Waters, G. and H. Hempel (Eds.), Wirtschaftsverlag NW, Bremerhaven (pp. 17-36). https://epic.awi.de/id/eprint/23118/

<!-- Food web studies about Weddell Sea / Key species or species role -->
The network complexity of the Weddell Sea food web is high comprising 488 species and 16200 predator-prey interactions [@Jacob2011]. In an attempt to better understand species roles related to food web stability, Jacob et al. (2011) performed secondary extinction experiments and found that the removal of small to medium-sized, and not large, organisms caused a cascade of secondary extinctions. This findings highlighted the relative importance of predators, rather than prey, for the architecture, functioning and stability of the Weddell Sea food web, which coincides with findings from recent meta-analyses in natural complex food webs [@Brose2019; @Perkins2022]. Other investigations considered this food web in a meta-analysis context showing that high predator-prey body-mass ratios are found for predator groups with specific trait combinations, including small vertebrates and large swimming or flying predators [@Brose2019]. These trait combinations generate weak interactions that stabilize communities against perturbations maintaining ecosystem functioning.

<!-- Objectives -->
In this study, we aim to go beyond a purely topological (presence/absence) assessment of who eats whom in the Weddell Sea ecosystem by providing a quantitative analysis of the trophic interaction network. We aim to analyse the species' role for the structure and stability of the food web. To achieve this, we: 1) estimated the strength for each interaction in the Weddell Sea food web, 2) characterised species' role considering both weighted and unweighted properties, and 3) analysed the species’ role related to the stability of the food web. This is the first time that interaction strengths were estimated for all pairwise trophic interactions at the species level (except for a few) for the Weddell sea food web.

@@

The high Antarctic Weddell Sea shelf is situated between 74 and 78ºS, stretching approximately 450 km from East to West (Figure 1). Water depth varies between 200 and 500 meters, and shallower areas are covered by continental ice, which forms the coastline along the eastern and southern part of the Weddell Sea. The shelf area contains a complex three-dimensional benthic habitat with large benthic biomasses, intermediate to high diversity in comparison to benthic boreal communities and a spatially patchy distribution of organisms [@Dayton1990; @Teixido2002].

@@

To analyse the impact of species on food web stability, we performed extinction simulations deleting one species at a time, that is for every extinction, network size was reduced by one species only. After each extinction, we calculated the stability of the network minus the targeted species (489 nodes) and compared it with that of the whole network (490 nodes in total). To calculate stability, we used the mean of the real part of the maximum eigenvalue of the Jacobian matrix using randomized Jacobians, keeping the predator-prey sign structure fixed [@Allesina2008; @Grilli2016]. This stability index indicates a more stable food web when it is negative. We performed 1000 simulations for each species removal and obtained a mean maximum eigenvalue for each case. Finally, we statistically analysed this difference with an Anderson-Darling test considering a p-value < 0.01 [@Scholz1987]. If the difference between the networks without and with the removed species is positive, then the stability of the food web is higher without the targeted species, and vice versa. A detailed description on the stability calculations can be found in the supplementary material.

To identify the species with a relatively high effect on food web stability, we plotted the results for each species extinction on food web stability, considering weighted (interaction strength) and unweighted properties (WHICH?), and species habitat.

@@

We employed a range of descriptors using both unweighted and weighted food web properties to characterise the dynamic and multifaceted nature of the Weddell Sea food web. Our results show a positive relationship between interaction strength and trophic level, and between interaction strength and species degree. In the Weddell Sea, species with high degree also tend to have high mean interaction strengths. This positive relationship between interaction strength and species degree reinforces the central role of species with many interactions: species with a high degree (hubs) have a large impact on overall food web structure and functioning [@Dunne2002a; @Kortsch2015]. On the other hand, the positive interaction strength-trophic level relationship contradicts studies that suggest that mid-trophic level species (e.g. krill, mesopelagic fish, squid) are involved in the major pathways of energy flow in high-latitude marine ecosystems [@Pinkerton2014; @Murphy2016; @McCormack2020; @Riccialdelli2020]. Such contradiction could be explained by the lack of species biomass information in the calculation of interaction strength we applied here [@Pawar2012]. Although this methodology allows information on species biomass or density to be included, this type of data was not available for the majority of species of the Weddell Sea food web.

Overall, the combination of information on the quantity and quality of interactions and its relationship enables a robust assessment of the species’ role in the stability of the food web [@Cirtwill2018a].

@@

Only a few species play a key role with respect to the Weddell Sea food web stability, according to the stability index employed in this study. This is in concordance with other studies on complex empirical food webs in marine ecosystems in the Arctic and other locations in Antarctica [@Kortsch2015; @Marina2018; @Rodriguez2022]. These key species are characterised by a particular set of food web properties: high to mean IS; mid to high trophic level; a relatively high number of interactions; and mid to low trophic similarity. In a previous study on sequential extinction simulations for the Weddell Sea food web [@Jacob2011], it was found  that larger bodied-sized species could be lost without causing a collapse of the network. A major caveat of this finding, also recognised by the authors, was that population dynamics were ignored and hence no top-down extinctions, or other indirect effects, could occur. In our study we considered such top-down effects by including information on the species IS, which is of paramount importance when analysing the response of perturbations in ecological communities [@McCann1998; @Montoya2009; @Novak2011]. Thus, our study suggests that species with high mean IS and high trophic levels need to be considered with particular attention when trying to predict the effects of perturbations on the Weddell Sea ecosystem. This conclusion is further reinforced by the finding that these species have mid to low trophic similarity, which means that few other species of the food web can occupy the same trophic role. In a review, it was emphasised that polar pelagic communities are particularly sensitive to changes due to a low functional redundancy at key trophic levels [@Murphy2016]. Here we provide a broader analysis of the species impact on food web robustness by including species from all habitats (benthic, pelagic and land-based). This suggests that the sensitivity of marine polar ecosystems to environmental perturbations is a concern also beyond the pelagic realm.

@@

We consider that the information provided in this study is important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea region [@Teschke2021].
